---
title: "Cardio — Analysis"
author: "Tianqi Ouyang"
date: last-modified
format:
  html:
    echo: true
    code-fold: true
    toc: true
    toc-depth: 3
---

```{r setup}
#| include: false
knitr::opts_chunk$set(warning = FALSE, message = FALSE)

# Source the shared pipeline. Pass output = tempfile() so we don't leave a
# stray shared_pipeline.R behind, and parallel renders don't race on the path.
source(knitr::purl(
  "/Users/to909/Desktop/Terlipressin projects/Terlipressin/code/shared_pipeline.qmd",
  output = tempfile(fileext = ".R"),
  quiet  = TRUE
))

library(tidyverse)
library(tableone)
library(splines)
library(ggplot2)
```

# Cohorts

Two cohorts are used.

**Echo cohort** — at least one of the five core echo measurements is available
(EF, lateral e′, GLS, E/A, E/e′):

```{r echo-cohort}
#| code-summary: "Show echo-cohort filter"
echo_cohort <- master %>%
  filter(
    echo_ef_notavailable___999      == 0 |
    echo_late_notavailable___999    == 0 |
    echo_gls_notavailable___999     == 0 |
    echo_earatio_notavailable___999 == 0 |
    echo_eeratio_notavailable___999 == 0
  )

cat("Echo cohort N:", nrow(echo_cohort), "\n")
```

**Pre-terli echo cohort** — echo cohort AND the echo was done before terlipressin
initiation (`days_to_echo < days_to_terli_initiation_day0`). Used for tasks 4–8.

```{r preterli-cohort}
#| code-summary: "Show pre-terli echo cohort filter"
preterli_echo <- echo_cohort %>%
  filter(days_to_echo < days_to_terli_initiation_day0)

cat("Pre-terli echo cohort N:", nrow(preterli_echo), "\n")
```

# Tasks 1–3 — Echo Timing Counts

Counts within the echo cohort, split by whether the echo was before, on the same
day as, or after terlipressin initiation.

```{r timing-counts}
#| code-summary: "Show timing-count code"
timing_counts <- echo_cohort %>%
  mutate(timing = case_when(
    days_to_echo <  days_to_terli_initiation_day0 ~ "1. echo before terli",
    days_to_echo == days_to_terli_initiation_day0 ~ "2. echo same day as terli",
    days_to_echo >  days_to_terli_initiation_day0 ~ "3. echo after terli",
    TRUE                                          ~ "missing"
  )) %>%
  count(timing, name = "N")

knitr::kable(timing_counts, caption = "Patients in echo cohort by echo timing relative to terlipressin initiation")
```

# Task 4 — Demographics Table (stratified by `hrs_responders`)

```{r task4-demographics}
#| code-summary: "Show demographic variable lists + Table 1 call"
demographic_categorical <- c(
  "sex_male", "race", "hispanic",
  "cirrhosis_etiology", "etoh_hepatitis",
  "hegrade_admit", "admit_ascites",
  "htn_medical_hx", "dm_medical_hx", "ckd_medical_hx", "cad_medical_hx",
  "sbp_ppx_preadmission", "lactulose_preadmission",
  "ascites_previous_comp", "refascites_previous_comp",
  "he_previous_comp", "sbp_previous_comp", "nonsbp_previous_comp",
  "phtn_bleed_previous_comp", "hcc_previous_comp", "tips_previous_comp",
  "baseline_scr_outpt", "aki_previous_90days", "transfer_aki",
  "aki_precipitant_nsbpinf",
  "he_aki", "rrt_aki", "mv_aki", "pressors_aki",
  "chest_imaging", "chest_imaging_pulmedema", "chest_imaging_effusion",
  "ecg", "echo",
  "albumin_before_terli", "crystalloid_before_terli",
  "nsbb_before_terli", "mido_before_terli", "oct_before_terli", "ne_before_terli",
  "gihep_consult", "neph_consult", "pall_consult",
  "code_status",
  "ua_hrs", "ua_na_hrs", "ua_cr_hrs", "ua_urea_hrs",
  "ua_granularcasts_hrs", "ua_epi_cells_hrs",
  "location_terli_day0", "terli_route_day0",
  "tele_terli_day0", "he_terli_day0",
  "infection_terli_day0", "typeinfection_terli_day0",
  "aclf_grade", "akin"
)

demographic_numerical <- c("age")

# Keep only variables that actually exist in the dataset
demographic_categorical <- intersect(demographic_categorical, names(preterli_echo))
demographic_numerical   <- intersect(demographic_numerical,   names(preterli_echo))
demographic_all <- c(demographic_numerical, demographic_categorical)

dat4 <- preterli_echo %>%
  mutate(across(all_of(demographic_categorical), as.factor),
         across(all_of(demographic_numerical),   as.numeric))

create_table_one(
  all_vars = demographic_all,
  cat_vars = demographic_categorical,
  data     = dat4,
  strata   = "hrs_responders",
  output   = "../Results/table_demographics_hrs_responders.csv"
)
```

Output: [`Results/table_demographics_hrs_responders.csv`](../Results/table_demographics_hrs_responders.csv)

::: {.callout-warning}
## `echo_gls` quirk

In the pre-terli echo cohort, `echo_gls` ranges from **−20 to +25** — some centers store GLS as a negative number (the standard convention) and others store the absolute value. **83 %** of the cohort is missing GLS. Tasks 5–7 below report raw values only; medians are unstable and any future modeling on GLS should harmonize sign per center or work with `abs(echo_gls)`.
:::

# Task 5 — HRS Response (binary) × Echo Continuous Values

```{r task5-response}
#| code-summary: "Show Task 5 Table 1 call"
echo_continuous <- c("echo_ef", "echo_earatio", "echo_eeratio",
                     "echo_gls", "echo_lateral_e")

dat5 <- preterli_echo %>%
  mutate(across(all_of(echo_continuous), as.numeric))

create_table_one(
  all_vars = echo_continuous,
  cat_vars = NULL,                       # all numerical
  data     = dat5,
  strata   = "hrs_responders_cat_2",
  output   = "../Results/table_echo_hrs_responders_cat_2.csv"
)
```

Output: [`Results/table_echo_hrs_responders_cat_2.csv`](../Results/table_echo_hrs_responders_cat_2.csv)

# Task 6 — Respiratory Failure × Echo Continuous Values

```{r task6-respfail}
#| code-summary: "Show Task 6 Table 1 call"
dat6 <- preterli_echo %>%
  mutate(across(all_of(echo_continuous), as.numeric))

create_table_one(
  all_vars = echo_continuous,
  cat_vars = NULL,
  data     = dat6,
  strata   = "terli_ae_respfail",
  output   = "../Results/table_echo_terli_ae_respfail.csv"
)
```

Output: [`Results/table_echo_terli_ae_respfail.csv`](../Results/table_echo_terli_ae_respfail.csv)

# Task 7 — 90-Day Mortality × Echo Continuous Values

Uses `time_to_death_status_90days` (death capped at 90 days).

```{r task7-mortality}
#| code-summary: "Show Task 7 Table 1 call"
dat7 <- preterli_echo %>%
  mutate(across(all_of(echo_continuous), as.numeric))

create_table_one(
  all_vars = echo_continuous,
  cat_vars = NULL,
  data     = dat7,
  strata   = "time_to_death_status_90days",
  output   = "../Results/table_echo_time_to_death_status_90days.csv"
)
```

Output: [`Results/table_echo_time_to_death_status_90days.csv`](../Results/table_echo_time_to_death_status_90days.csv)

# Task 8 — Cubic-Spline Logistic Regression

Outcome: `hrs_responders_cat_2` (binary). Covariate: `echo_ef` modelled as a
natural cubic spline with 3 degrees of freedom.

```{r task8-spline-fit}
#| code-summary: "Show spline model fit"
dat8 <- preterli_echo %>%
  mutate(
    hrs_responders_cat_2 = as.numeric(as.character(hrs_responders_cat_2)),
    echo_ef              = as.numeric(echo_ef)
  ) %>%
  drop_na(echo_ef, hrs_responders_cat_2)

fit8 <- glm(
  hrs_responders_cat_2 ~ ns(echo_ef, df = 3),
  data   = dat8,
  family = binomial
)

summary(fit8)
```

```{r task8-spline-plot, fig.height=5, fig.width=7}
#| code-summary: "Show predicted-probability plot"
grid8 <- tibble(
  echo_ef = seq(min(dat8$echo_ef, na.rm = TRUE),
                max(dat8$echo_ef, na.rm = TRUE),
                length.out = 200)
)

pred8 <- predict(fit8, newdata = grid8, type = "link", se.fit = TRUE)
grid8 <- grid8 %>%
  mutate(
    fit   = plogis(pred8$fit),
    lower = plogis(pred8$fit - 1.96 * pred8$se.fit),
    upper = plogis(pred8$fit + 1.96 * pred8$se.fit)
  )

ggplot(grid8, aes(echo_ef, fit)) +
  geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.2, fill = "#2c8dbb") +
  geom_line(linewidth = 1, color = "#2c8dbb") +
  geom_rug(data = dat8, aes(x = echo_ef), inherit.aes = FALSE,
           alpha = 0.3, sides = "b") +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1),
                     limits = c(0, 1)) +
  labs(
    x = "Ejection fraction (%)",
    y = "Probability of HRS response (hrs_responders_cat_2 = 1)",
    title = "Predicted probability of HRS response by EF (natural cubic spline, df = 3)",
    caption = "Shaded ribbon = 95% CI; rug = observed EF values"
  ) +
  theme_bw(base_size = 12)
```
