---
title: "Cardio — Analysis"
author: "Tianqi Ouyang"
date: last-modified
format:
  html:
    echo: true
    code-fold: true
    toc: true
    toc-depth: 3
---

```{r setup}
#| include: false
knitr::opts_chunk$set(warning = FALSE, message = FALSE)

# Load `master` and shared helpers from the main project.
source(knitr::purl(
  "/Users/to909/Desktop/Terlipressin projects/Terlipressin/code/shared_pipeline.qmd",
  output = tempfile(fileext = ".R"),
  quiet  = TRUE
))

library(tidyverse)
library(tableone)
library(splines)
library(ggplot2)

# Make sure the per-page download dir exists.
dir.create("tables", showWarnings = FALSE)

# --- Local override of `create_table_one()` ---------------------------------
# Same behaviour as the analyst-pasted version — only writes a CSV. The single
# change vs. the pasted snippet is the trailing `invisible(T1_after_print)` so
# the inline kable below each chunk can render the same table the CSV holds.
create_table_one <- function(all_vars,
                             data,
                             strata,
                             output,
                             cat_vars = NULL,
                             exact = NULL) {

  if (is.null(cat_vars)) {
    numerical_vars <- all_vars
  } else {
    numerical_vars <- setdiff(all_vars, cat_vars)
  }

  T1_after <- tableone::CreateTableOne(
    vars       = all_vars,
    strata     = strata,
    data       = data,
    factorVars = cat_vars,
    addOverall = TRUE,
    includeNA  = FALSE
  )

  T1_after_print <- print(
    T1_after,
    showAllLevels = TRUE,
    exact         = exact,
    nonnormal     = numerical_vars,
    missing       = TRUE,
    quote         = FALSE,
    noSpaces      = TRUE,
    printToggle   = FALSE
  )

  write.csv(T1_after_print, file = output, row.names = TRUE)
  invisible(T1_after_print)
}
```

# Cohorts

```{r cohorts}
#| code-summary: "Show cohort filters"
echo_cohort <- master %>%
  filter(
    echo_ef_notavailable___999      == 0 |
    echo_late_notavailable___999    == 0 |
    echo_gls_notavailable___999     == 0 |
    echo_earatio_notavailable___999 == 0 |
    echo_eeratio_notavailable___999 == 0
  )

preterli_echo <- echo_cohort %>%
  filter(days_to_echo < days_to_terli_initiation_day0) %>%
  mutate(
    # Abnormal-echo flags (1 = abnormal, 0 = normal, NA if measurement missing)
    echo_ef_abnormal        = if_else(is.na(echo_ef),
                                      NA_real_, as.numeric(echo_ef < 60)),
    echo_eeratio_abnormal   = if_else(is.na(echo_eeratio),
                                      NA_real_, as.numeric(echo_eeratio > 15)),
    # GLS uses |GLS| < 18 because some centers store the value as a negative
    # number (standard convention) while others store the absolute value.
    echo_gls_abnormal       = if_else(is.na(echo_gls),
                                      NA_real_, as.numeric(abs(echo_gls) < 18)),
    echo_lateral_e_abnormal = if_else(is.na(echo_lateral_e),
                                      NA_real_, as.numeric(echo_lateral_e < 10)),
    echo_tr_abnormal        = if_else(is.na(echo_tr),
                                      NA_real_, as.numeric(echo_tr > 2.8)),
    echo_lavi_abnormal      = if_else(is.na(echo_lavi),
                                      NA_real_, as.numeric(echo_lavi > 34))
  )

cat("Echo cohort N      :", nrow(echo_cohort),    "\n")
cat("Pre-terli cohort N :", nrow(preterli_echo), "\n")
```

# Tasks 1–3 — Echo Timing Counts

```{r timing-counts}
#| code-summary: "Show timing-count code"
timing_counts <- echo_cohort %>%
  mutate(timing = case_when(
    days_to_echo <  days_to_terli_initiation_day0 ~ "1. echo before terli",
    days_to_echo == days_to_terli_initiation_day0 ~ "2. echo same day as terli",
    days_to_echo >  days_to_terli_initiation_day0 ~ "3. echo after terli",
    TRUE                                          ~ "missing"
  )) %>%
  count(timing, name = "N")

knitr::kable(timing_counts,
             caption = "Patients in echo cohort by echo timing relative to terlipressin initiation")
```

# Task 4 — Demographics × `hrs_responders`

```{r task4-demographics}
#| code-summary: "Show demographic variable lists + Table 1 call"
demographic_categorical <- c(
  "sex_male", "race", "hispanic",
  "cirrhosis_etiology", "etoh_hepatitis",
  "hegrade_admit", "admit_ascites",
  "htn_medical_hx", "dm_medical_hx", "ckd_medical_hx", "cad_medical_hx",
  "sbp_ppx_preadmission", "lactulose_preadmission",
  "ascites_previous_comp", "refascites_previous_comp",
  "he_previous_comp", "sbp_previous_comp", "nonsbp_previous_comp",
  "phtn_bleed_previous_comp", "hcc_previous_comp", "tips_previous_comp",
  "baseline_scr_outpt", "aki_previous_90days", "transfer_aki",
  "aki_precipitant_nsbpinf",
  "he_aki", "rrt_aki", "mv_aki", "pressors_aki",
  "chest_imaging", "chest_imaging_pulmedema", "chest_imaging_effusion",
  "ecg", "echo",
  "albumin_before_terli", "crystalloid_before_terli",
  "nsbb_before_terli", "mido_before_terli", "oct_before_terli", "ne_before_terli",
  "gihep_consult", "neph_consult", "pall_consult",
  "code_status",
  "ua_hrs", "ua_na_hrs", "ua_cr_hrs", "ua_urea_hrs",
  "ua_granularcasts_hrs", "ua_epi_cells_hrs",
  "location_terli_day0", "terli_route_day0",
  "tele_terli_day0", "he_terli_day0",
  "infection_terli_day0", "typeinfection_terli_day0",
  "aclf_grade", "akin"
)
demographic_numerical <- c("age")

demographic_categorical <- intersect(demographic_categorical, names(preterli_echo))
demographic_numerical   <- intersect(demographic_numerical,   names(preterli_echo))
demographic_all <- c(demographic_numerical, demographic_categorical)

dat4 <- preterli_echo %>%
  mutate(across(all_of(demographic_categorical), as.factor),
         across(all_of(demographic_numerical),   as.numeric))

t4 <- create_table_one(
  all_vars = demographic_all,
  cat_vars = demographic_categorical,
  data     = dat4,
  strata   = "hrs_responders",
  output   = "tables/task4_demographics_hrs_responders.csv"
)
knitr::kable(t4, caption = "Task 4 — Demographics by `hrs_responders` (3-level)")
```

[📥 Download CSV — Task 4](tables/task4_demographics_hrs_responders.csv)

::: {.callout-warning}
## `echo_gls` quirk

In the pre-terli echo cohort, `echo_gls` ranges from **−20 to +25** — some centers store GLS as a negative number (the standard convention) and others store the absolute value. **83 %** of the cohort is missing GLS. Tasks 5–7 below report raw values only; medians are unstable and any future modeling on GLS should harmonize sign per center or work with `abs(echo_gls)`.
:::

# Task 5 — Echo × `hrs_responders_cat_2`

```{r task5-response}
#| code-summary: "Show Task 5 Table 1 call"
echo_continuous <- c("echo_ef", "echo_earatio", "echo_eeratio",
                     "echo_gls", "echo_lateral_e")
echo_flags <- c("echo_ef_abnormal", "echo_eeratio_abnormal", "echo_gls_abnormal",
                "echo_lateral_e_abnormal", "echo_tr_abnormal", "echo_lavi_abnormal")

dat5 <- preterli_echo %>%
  mutate(across(all_of(echo_continuous), as.numeric),
         across(all_of(echo_flags),      as.factor))

t5 <- create_table_one(
  all_vars = c(echo_continuous, echo_flags),
  cat_vars = echo_flags,
  data     = dat5,
  strata   = "hrs_responders_cat_2",
  output   = "tables/task5_echo_hrs_responders_cat_2.csv"
)
knitr::kable(t5, caption = "Task 5 — Echo by HRS response (binary)")
```

[📥 Download CSV — Task 5](tables/task5_echo_hrs_responders_cat_2.csv)

# Task 6 — Echo × `terli_ae_respfail`

```{r task6-respfail}
#| code-summary: "Show Task 6 Table 1 call"
dat6 <- preterli_echo %>%
  mutate(across(all_of(echo_continuous), as.numeric),
         across(all_of(echo_flags),      as.factor))

t6 <- create_table_one(
  all_vars = c(echo_continuous, echo_flags),
  cat_vars = echo_flags,
  data     = dat6,
  strata   = "terli_ae_respfail",
  output   = "tables/task6_echo_terli_ae_respfail.csv"
)
knitr::kable(t6, caption = "Task 6 — Echo by terlipressin-associated respiratory failure")
```

[📥 Download CSV — Task 6](tables/task6_echo_terli_ae_respfail.csv)

# Task 7 — Echo × `time_to_death_status_90days`

Death capped at 90 days.

```{r task7-mortality}
#| code-summary: "Show Task 7 Table 1 call"
dat7 <- preterli_echo %>%
  mutate(across(all_of(echo_continuous), as.numeric),
         across(all_of(echo_flags),      as.factor))

t7 <- create_table_one(
  all_vars = c(echo_continuous, echo_flags),
  cat_vars = echo_flags,
  data     = dat7,
  strata   = "time_to_death_status_90days",
  output   = "tables/task7_echo_time_to_death_status_90days.csv"
)
knitr::kable(t7, caption = "Task 7 — Echo by 90-day mortality status")
```

[📥 Download CSV — Task 7](tables/task7_echo_time_to_death_status_90days.csv)

# Task 8 — Cubic-Spline Logistic Regression

Outcome: `hrs_responders_cat_2` (binary). Covariate: `echo_ef` modelled as a
natural cubic spline with 3 degrees of freedom.

```{r task8-spline-fit}
#| code-summary: "Show spline model fit"
dat8 <- preterli_echo %>%
  mutate(
    hrs_responders_cat_2 = as.numeric(as.character(hrs_responders_cat_2)),
    echo_ef              = as.numeric(echo_ef)
  ) %>%
  drop_na(echo_ef, hrs_responders_cat_2)

fit8 <- glm(
  hrs_responders_cat_2 ~ ns(echo_ef, df = 3),
  data   = dat8,
  family = binomial
)

summary(fit8)
```

```{r task8-spline-plot, fig.height=5, fig.width=7}
#| code-summary: "Show predicted-probability plot"
grid8 <- tibble(
  echo_ef = seq(min(dat8$echo_ef, na.rm = TRUE),
                max(dat8$echo_ef, na.rm = TRUE),
                length.out = 200)
)

pred8 <- predict(fit8, newdata = grid8, type = "link", se.fit = TRUE)
grid8 <- grid8 %>%
  mutate(
    fit   = plogis(pred8$fit),
    lower = plogis(pred8$fit - 1.96 * pred8$se.fit),
    upper = plogis(pred8$fit + 1.96 * pred8$se.fit)
  )

ggplot(grid8, aes(echo_ef, fit)) +
  geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.2, fill = "#2c8dbb") +
  geom_line(linewidth = 1, color = "#2c8dbb") +
  geom_rug(data = dat8, aes(x = echo_ef), inherit.aes = FALSE,
           alpha = 0.3, sides = "b") +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1),
                     limits = c(0, 1)) +
  labs(
    x = "Ejection fraction (%)",
    y = "Probability of HRS response (hrs_responders_cat_2 = 1)",
    title = "Predicted probability of HRS response by EF (natural cubic spline, df = 3)",
    caption = "Shaded ribbon = 95% CI; rug = observed EF values"
  ) +
  theme_bw(base_size = 12)
```
