---
title: "Cardio — Analysis"
author: "Tianqi Ouyang"
date: last-modified
format:
  html:
    echo: true
    code-fold: true
    toc: true
    toc-depth: 3
---

```{r setup}
#| include: false
knitr::opts_chunk$set(warning = FALSE, message = FALSE)

# Source the shared pipeline — loads `master` with all standard derived
# variables and exposes helper functions (including create_table_one).
source(knitr::purl(
  "/Users/to909/Desktop/Terlipressin projects/Terlipressin/code/shared_pipeline.qmd",
  quiet = TRUE
))

library(tidyverse)
library(tableone)
library(gtsummary)
library(survival)
library(survminer)
library(ggsurvfit)
library(tidycmprsk)
```

# Cohort & Derived Variables

Restrict to patients with at least one non-missing echo measurement among EF, E/A, E/e′, GLS, and lateral e′. Then derive six binary abnormal-echo flags.

```{r cohort-derive}
#| code-summary: "Show cohort filter + abnormal-echo flag derivation"
final_master <- master %>%
  filter(if_any(
    c(echo_ef, echo_earatio, echo_eeratio, echo_gls, echo_lateral_e),
    ~ !is.na(.)
  )) %>%
  mutate(
    echo_ef_abnormal       = if_else(is.na(echo_ef),         NA_real_, if_else(echo_ef        <= 50, 1, 0)),
    echo_eeratio_abnormal  = if_else(is.na(echo_eeratio),    NA_real_, if_else(echo_eeratio   >= 15, 1, 0)),
    echo_gls_abnormal      = if_else(is.na(echo_gls),        NA_real_, if_else(echo_gls       <  18, 1, 0)),
    echo_lateral_e_abnormal= if_else(is.na(echo_lateral_e),  NA_real_, if_else(echo_lateral_e <  10, 1, 0)),
    echo_tr_abnormal       = if_else(is.na(echo_tr),         NA_real_, if_else(echo_tr        > 2.8, 1, 0)),
    echo_lavi_abnormal     = if_else(is.na(echo_lavi),       NA_real_, if_else(echo_lavi      >  34, 1, 0))
  )

cat("Cohort size (any non-NA echo):", nrow(final_master), "\n")
```

# Variable Lists

```{r var-lists}
#| code-summary: "Show numeric and categorical echo variable lists"
num_vars_echo <- c("echo_ef", "echo_earatio", "echo_eeratio",
                   "echo_gls", "echo_lateral_e")

abnormal_flags <- c("echo_ef_abnormal", "echo_eeratio_abnormal",
                    "echo_gls_abnormal", "echo_lateral_e_abnormal",
                    "echo_tr_abnormal",  "echo_lavi_abnormal")

all_vars_echo <- unique(c(num_vars_echo, abnormal_flags))
cat_vars_echo <- setdiff(all_vars_echo, num_vars_echo)
```

# Table 1 — Stratified by HRS Responders (3-level)

```{r t1-hrs-responders}
#| code-summary: "Show Table 1 code"
create_table_one(
  all_vars = all_vars_echo,
  cat_vars = cat_vars_echo,
  data     = final_master,
  strata   = "hrs_responders",
  output   = "../Results/table_hrs_responders_a.csv"
)
```

# Table 1 — Stratified by HRS Responders (Binary)

```{r t1-hrs-cat2}
#| code-summary: "Show Table 1 code"
create_table_one(
  all_vars = all_vars_echo,
  cat_vars = cat_vars_echo,
  data     = final_master,
  strata   = "hrs_responders_cat_2",
  output   = "../Results/table_hrs_responders_cat_2_b.csv"
)
```

# Table 1 — Stratified by 90-Day Mortality

```{r t1-death-90d}
#| code-summary: "Show Table 1 code"
create_table_one(
  all_vars = all_vars_echo,
  cat_vars = cat_vars_echo,
  data     = final_master,
  strata   = "time_to_death_status_90days",
  output   = "../Results/table_time_to_death_status_90days.csv"
)
```

# Table 1 — Stratified by Terlipressin Respiratory Failure

Fisher exact used for every categorical row (`exact = cat_vars_echo`) because the respiratory-failure stratum is sparse.

```{r t1-respfail}
#| code-summary: "Show Table 1 code"
create_table_one(
  all_vars = all_vars_echo,
  cat_vars = cat_vars_echo,
  exact    = cat_vars_echo,
  data     = final_master,
  strata   = "terli_ae_respfail",
  output   = "../Results/table_terli_ae_respfail.csv"
)
```

# Availability of Echo Measurements

```{r echo-availability}
#| code-summary: "Show availability summary"
final_master %>%
  summarise(across(all_of(num_vars_echo), ~ sum(!is.na(.)))) %>%
  pivot_longer(everything(), names_to = "Variable", values_to = "N non-missing") %>%
  knitr::kable(caption = "Non-missing counts by echo measurement (cohort)")
```
