---
title: "Cardio — Analysis"
author: "Tianqi Ouyang"
date: last-modified
format:
  html:
    echo: true
    code-fold: true
    toc: true
    toc-depth: 3
---

```{r setup}
#| include: false
knitr::opts_chunk$set(warning = FALSE, message = FALSE)

# Load `master` and shared helpers from the main project.
source(knitr::purl(
  "/Users/to909/Desktop/Terlipressin projects/Terlipressin/code/shared_pipeline.qmd",
  output = tempfile(fileext = ".R"),
  quiet  = TRUE
))

library(tidyverse)
library(tableone)
library(splines)
library(ggplot2)
library(pROC)

# Make sure the per-page download dir exists.
dir.create("tables", showWarnings = FALSE)

# --- Local override of `create_table_one()` ---------------------------------
# Same behaviour as the analyst-pasted version — only writes a CSV. The single
# change vs. the pasted snippet is the trailing `invisible(T1_after_print)` so
# the inline kable below each chunk can render the same table the CSV holds.
create_table_one <- function(all_vars,
                             data,
                             strata,
                             output,
                             cat_vars = NULL,
                             exact = NULL) {

  if (is.null(cat_vars)) {
    numerical_vars <- all_vars
  } else {
    numerical_vars <- setdiff(all_vars, cat_vars)
  }

  T1_after <- tableone::CreateTableOne(
    vars       = all_vars,
    strata     = strata,
    data       = data,
    factorVars = cat_vars,
    addOverall = TRUE,
    includeNA  = FALSE
  )

  T1_after_print <- print(
    T1_after,
    showAllLevels = TRUE,
    exact         = exact,
    nonnormal     = numerical_vars,
    missing       = TRUE,
    quote         = FALSE,
    noSpaces      = TRUE,
    printToggle   = FALSE
  )

  write.csv(T1_after_print, file = output, row.names = TRUE)
  invisible(T1_after_print)
}

# --- Reusable spline logistic-regression helper ----------------------------
# Same parameters as the original Task 8 fit (natural cubic spline, df = 3,
# outcome = hrs_responders_cat_2, 95% CI ribbon, rug of observed x values).
spline_logit <- function(data, xvar,
                         outcome      = "hrs_responders_cat_2",
                         df           = 3,
                         xlab         = xvar,
                         cohort_label = "") {
  d <- data %>%
    mutate(
      .x = as.numeric(.data[[xvar]]),
      .y = as.numeric(as.character(.data[[outcome]]))
    ) %>%
    drop_na(.x, .y)

  fit <- glm(.y ~ ns(.x, df = df), data = d, family = binomial)

  grid <- tibble(.x = seq(min(d$.x), max(d$.x), length.out = 200))
  pred <- predict(fit, newdata = grid, type = "link", se.fit = TRUE)
  grid <- grid %>%
    mutate(
      fit   = plogis(pred$fit),
      lower = plogis(pred$fit - 1.96 * pred$se.fit),
      upper = plogis(pred$fit + 1.96 * pred$se.fit)
    )

  plot <- ggplot(grid, aes(.x, fit)) +
    geom_ribbon(aes(ymin = lower, ymax = upper),
                alpha = 0.2, fill = "#2c8dbb") +
    geom_line(linewidth = 1, color = "#2c8dbb") +
    geom_rug(data = d, aes(x = .x), inherit.aes = FALSE,
             alpha = 0.3, sides = "b") +
    scale_y_continuous(labels = scales::percent_format(accuracy = 1),
                       limits = c(0, 1)) +
    labs(
      x       = xlab,
      y       = sprintf("Probability of HRS response (%s = 1)", outcome),
      title   = sprintf("P(%s = 1) by %s — %s", outcome, xvar, cohort_label),
      caption = sprintf("Natural cubic spline, df = %d; ribbon = 95%% CI; rug = observed (n = %d)",
                        df, nrow(d))
    ) +
    theme_bw(base_size = 12)

  list(fit = fit, plot = plot, n = nrow(d))
}
```

# Cohorts

```{r cohorts}
#| code-summary: "Show cohort filters"
echo_cohort <- master %>%
  filter(
    echo_ef_notavailable___999      == 0 |
    echo_late_notavailable___999    == 0 |
    echo_gls_notavailable___999     == 0 |
    echo_earatio_notavailable___999 == 0 |
    echo_eeratio_notavailable___999 == 0
  ) %>%
  # Harmonize echo_gls sign convention: some centers store negatives (the
  # standard convention), others store the absolute value. Take abs() once
  # here so every downstream summary (medians, abnormal flag, spline) uses
  # magnitudes only.
  mutate(echo_gls = abs(echo_gls))

preterli_echo_base <- echo_cohort %>%
  # Pre-terli OR same-day echo (per analyst spec — adds 13 same-day patients
  # on top of the strict pre-terli set).
  filter(days_to_echo <= days_to_terli_initiation_day0)

# --- Youden-optimal echo_ef cutoff (Task 9 input) --------------------------
# Compute once on the cohort so the same cutoff is used to (a) derive the
# echo_ef_abnormal_youden flag below and (b) render Task 9.
.youden_data <- preterli_echo_base %>%
  filter(!is.na(echo_ef), !is.na(hrs_responders_cat_2))
.roc_ef <- pROC::roc(.youden_data$hrs_responders_cat_2, .youden_data$echo_ef,
                     quiet = TRUE, direction = "auto")
.youden_coords <- pROC::coords(.roc_ef, x = "best", best.method = "youden",
                               ret = c("threshold","sensitivity",
                                       "specificity","accuracy","ppv","npv"))
youden_ef_cutoff <- as.numeric(.youden_coords$threshold)

preterli_echo <- preterli_echo_base %>%
  mutate(
    # Abnormal-echo flags (1 = abnormal, 0 = normal, NA if measurement missing)
    echo_ef_abnormal        = if_else(is.na(echo_ef),
                                      NA_real_, as.numeric(echo_ef < 60)),
    echo_ef_abnormal_youden = if_else(is.na(echo_ef),
                                      NA_real_, as.numeric(echo_ef < youden_ef_cutoff)),
    echo_eeratio_abnormal   = if_else(is.na(echo_eeratio),
                                      NA_real_, as.numeric(echo_eeratio > 15)),
    # echo_gls is already stored as a magnitude (see harmonization in echo_cohort).
    echo_gls_abnormal       = if_else(is.na(echo_gls),
                                      NA_real_, as.numeric(echo_gls < 18)),
    echo_lateral_e_abnormal = if_else(is.na(echo_lateral_e),
                                      NA_real_, as.numeric(echo_lateral_e < 10)),
    # TR is unit-aware: < 100 treated as m/s (cutoff 2.8); >= 100 as cm/s (cutoff 280).
    echo_tr_abnormal        = if_else(is.na(echo_tr),
                                      NA_real_,
                                      as.numeric(if_else(echo_tr < 100,
                                                         echo_tr > 2.8,
                                                         echo_tr > 280))),
    echo_lavi_abnormal      = if_else(is.na(echo_lavi),
                                      NA_real_, as.numeric(echo_lavi > 34))
  )

cat("Echo cohort N      :", nrow(echo_cohort),    "\n")
cat("Pre-terli cohort N :", nrow(preterli_echo), "\n")
```

# Tasks 1–3 — Echo Timing Counts

**Cohort:** Echo cohort (N = `r nrow(echo_cohort)`)

```{r timing-counts}
#| code-summary: "Show timing-count code"
timing_counts <- echo_cohort %>%
  mutate(timing = case_when(
    days_to_echo <  days_to_terli_initiation_day0 ~ "1. echo before terli",
    days_to_echo == days_to_terli_initiation_day0 ~ "2. echo same day as terli",
    days_to_echo >  days_to_terli_initiation_day0 ~ "3. echo after terli",
    TRUE                                          ~ "missing"
  )) %>%
  count(timing, name = "N")

knitr::kable(timing_counts,
             caption = "Patients in echo cohort by echo timing relative to terlipressin initiation")
```

# Task 4 — Demographics × `hrs_responders`

**Cohort:** Pre-terli + same-day echo (N = `r nrow(preterli_echo)`)

```{r task4-demographics}
#| code-summary: "Show demographic variable lists + Table 1 call"
demographic_categorical <- c(
  "sex_male", "race", "hispanic",
  "cirrhosis_etiology", "etoh_hepatitis",
  "hegrade_admit", "admit_ascites",
  "htn_medical_hx", "dm_medical_hx", "ckd_medical_hx", "cad_medical_hx",
  "sbp_ppx_preadmission", "lactulose_preadmission",
  "ascites_previous_comp", "refascites_previous_comp",
  "he_previous_comp", "sbp_previous_comp", "nonsbp_previous_comp",
  "phtn_bleed_previous_comp", "hcc_previous_comp", "tips_previous_comp",
  "baseline_scr_outpt", "aki_previous_90days", "transfer_aki",
  "aki_precipitant_nsbpinf",
  "he_aki", "rrt_aki", "mv_aki", "pressors_aki",
  "chest_imaging", "chest_imaging_pulmedema", "chest_imaging_effusion",
  "ecg", "echo",
  "albumin_before_terli", "crystalloid_before_terli",
  "nsbb_before_terli", "mido_before_terli", "oct_before_terli", "ne_before_terli",
  "gihep_consult", "neph_consult", "pall_consult",
  "code_status",
  "ua_hrs", "ua_na_hrs", "ua_cr_hrs", "ua_urea_hrs",
  "ua_granularcasts_hrs", "ua_epi_cells_hrs",
  "location_terli_day0", "terli_route_day0",
  "tele_terli_day0", "he_terli_day0",
  "infection_terli_day0", "typeinfection_terli_day0",
  "aclf_grade", "akin"
)
demographic_numerical <- c("age")

demographic_categorical <- intersect(demographic_categorical, names(preterli_echo))
demographic_numerical   <- intersect(demographic_numerical,   names(preterli_echo))
demographic_all <- c(demographic_numerical, demographic_categorical)

dat4 <- preterli_echo %>%
  mutate(across(all_of(demographic_categorical), as.factor),
         across(all_of(demographic_numerical),   as.numeric))

t4 <- create_table_one(
  all_vars = demographic_all,
  cat_vars = demographic_categorical,
  data     = dat4,
  strata   = "hrs_responders",
  output   = "tables/task4_demographics_hrs_responders.csv"
)
knitr::kable(t4, caption = "Task 4 — Demographics by `hrs_responders` (3-level)")
```

[📥 Download CSV — Task 4](tables/task4_demographics_hrs_responders.csv)

# Task 5 — Echo × `hrs_responders_cat_2`

**Cohort:** Pre-terli + same-day echo (N = `r nrow(preterli_echo)`)

```{r task5-response}
#| code-summary: "Show Task 5 Table 1 call"
echo_continuous <- c("echo_ef", "echo_earatio", "echo_eeratio",
                     "echo_gls", "echo_lateral_e")
echo_flags <- c("echo_ef_abnormal", "echo_ef_abnormal_youden",
                "echo_eeratio_abnormal", "echo_gls_abnormal",
                "echo_lateral_e_abnormal", "echo_tr_abnormal", "echo_lavi_abnormal")

dat5 <- preterli_echo %>%
  mutate(across(all_of(echo_continuous), as.numeric),
         across(all_of(echo_flags),      as.factor))

t5 <- create_table_one(
  all_vars = c(echo_continuous, echo_flags),
  cat_vars = echo_flags,
  data     = dat5,
  strata   = "hrs_responders_cat_2",
  output   = "tables/task5_echo_hrs_responders_cat_2.csv"
)
knitr::kable(t5, caption = "Task 5 — Echo by HRS response (binary)")
```

[📥 Download CSV — Task 5](tables/task5_echo_hrs_responders_cat_2.csv)

# Task 6 — Echo × `terli_ae_respfail`

**Cohort:** Pre-terli + same-day echo (N = `r nrow(preterli_echo)`)

```{r task6-respfail}
#| code-summary: "Show Task 6 Table 1 call"
dat6 <- preterli_echo %>%
  mutate(across(all_of(echo_continuous), as.numeric),
         across(all_of(echo_flags),      as.factor))

t6 <- create_table_one(
  all_vars = c(echo_continuous, echo_flags),
  cat_vars = echo_flags,
  data     = dat6,
  strata   = "terli_ae_respfail",
  output   = "tables/task6_echo_terli_ae_respfail.csv"
)
knitr::kable(t6, caption = "Task 6 — Echo by terlipressin-associated respiratory failure")
```

[📥 Download CSV — Task 6](tables/task6_echo_terli_ae_respfail.csv)

# Task 7 — Echo × `time_to_death_status_90days`

**Cohort:** Pre-terli + same-day echo (N = `r nrow(preterli_echo)`)

Death capped at 90 days.

```{r task7-mortality}
#| code-summary: "Show Task 7 Table 1 call"
dat7 <- preterli_echo %>%
  mutate(across(all_of(echo_continuous), as.numeric),
         across(all_of(echo_flags),      as.factor))

t7 <- create_table_one(
  all_vars = c(echo_continuous, echo_flags),
  cat_vars = echo_flags,
  data     = dat7,
  strata   = "time_to_death_status_90days",
  output   = "tables/task7_echo_time_to_death_status_90days.csv"
)
knitr::kable(t7, caption = "Task 7 — Echo by 90-day mortality status")
```

[📥 Download CSV — Task 7](tables/task7_echo_time_to_death_status_90days.csv)

# Task 8 — Cubic-Spline Logistic Regression Models

Outcome: `hrs_responders_cat_2` (binary). Covariate `echo_ef` modelled as a
natural cubic spline with **df = 3**, fit on each cohort.

## 8.1 `echo_ef` — Pre-terli cohort

**Cohort:** Pre-terli + same-day echo (N = `r nrow(preterli_echo)`)

```{r task8-1, fig.height=5, fig.width=7}
#| code-summary: "Show spline fit + plot"
res_8_1 <- spline_logit(preterli_echo, "echo_ef",
                        xlab = "Ejection fraction (%)",
                        cohort_label = "Pre-terli cohort")
summary(res_8_1$fit)
res_8_1$plot
```

## 8.2 `echo_ef` — Echo cohort

**Cohort:** Echo cohort (N = `r nrow(echo_cohort)`)

```{r task8-2, fig.height=5, fig.width=7}
#| code-summary: "Show spline fit + plot"
res_8_2 <- spline_logit(echo_cohort, "echo_ef",
                        xlab = "Ejection fraction (%)",
                        cohort_label = "Echo cohort (all echos)")
summary(res_8_2$fit)
res_8_2$plot
```

# Task 9 — Optimal `echo_ef` Cutoff (Youden's J)

**Cohort:** Pre-terli + same-day echo (N = `r nrow(preterli_echo)`)

Find the EF threshold that maximises **Youden's J = sensitivity + specificity − 1**
for predicting HRS response (`hrs_responders_cat_2`). pROC chooses the direction
that maximises AUC. The resulting cutoff is used to derive the
**`echo_ef_abnormal_youden`** flag, which is now included as an additional row
in Tasks 5–7 alongside the clinical `echo_ef_abnormal` (EF < 60).

```{r task9-cutoff-table}
#| code-summary: "Show cutoff + AUC summary"
auc_ci <- as.numeric(pROC::ci.auc(.roc_ef))
tibble::tibble(
  Metric = c("Youden-optimal EF cutoff",
             "AUC", "AUC 95% CI (lower)", "AUC 95% CI (upper)",
             "Direction (pROC)",
             "Sensitivity at cutoff", "Specificity at cutoff",
             "Accuracy at cutoff", "PPV at cutoff", "NPV at cutoff",
             "N evaluable"),
  Value  = c(sprintf("%.1f", youden_ef_cutoff),
             sprintf("%.3f", as.numeric(pROC::auc(.roc_ef))),
             sprintf("%.3f", auc_ci[1]),
             sprintf("%.3f", auc_ci[3]),
             .roc_ef$direction,
             sprintf("%.3f", .youden_coords$sensitivity),
             sprintf("%.3f", .youden_coords$specificity),
             sprintf("%.3f", .youden_coords$accuracy),
             sprintf("%.3f", .youden_coords$ppv),
             sprintf("%.3f", .youden_coords$npv),
             as.character(nrow(.youden_data)))
) |>
  knitr::kable(caption = "Youden-optimal echo_ef cutoff for hrs_responders_cat_2")
```

::: {.callout-note}
## Direction note

pROC reports **direction = `r .roc_ef$direction`**, which in this cohort means
*higher* EF is associated with **non-response**. AUC is `r sprintf("%.3f", as.numeric(pROC::auc(.roc_ef)))` — weak discrimination, so this cutoff should be interpreted cautiously. The flag is defined as `echo_ef < `r sprintf("%.1f", youden_ef_cutoff)`` to preserve the clinical "abnormal = below cutoff" convention used by the existing `echo_ef_abnormal` (< 60).
:::

```{r task9-roc-plot, fig.height=5, fig.width=6}
#| code-summary: "Show ROC plot"
roc_df <- tibble::tibble(
  fpr = 1 - .roc_ef$specificities,
  tpr = .roc_ef$sensitivities
) |>
  arrange(fpr, tpr)

opt_point <- tibble::tibble(
  fpr = 1 - .youden_coords$specificity,
  tpr =     .youden_coords$sensitivity
)

ggplot(roc_df, aes(fpr, tpr)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "grey60") +
  geom_path(linewidth = 1, color = "#2c8dbb") +
  geom_point(data = opt_point, color = "#d73027", size = 3) +
  geom_text(data = opt_point,
            aes(label = sprintf("EF = %.1f\nsens = %.2f, spec = %.2f",
                                youden_ef_cutoff,
                                .youden_coords$sensitivity,
                                .youden_coords$specificity)),
            hjust = -0.05, vjust = 1.1, size = 3.5) +
  scale_x_continuous(limits = c(0, 1)) +
  scale_y_continuous(limits = c(0, 1.05)) +
  coord_equal() +
  labs(
    x = "1 − Specificity (false-positive rate)",
    y = "Sensitivity (true-positive rate)",
    title = sprintf("ROC of echo_ef for HRS response — AUC = %.3f",
                    as.numeric(pROC::auc(.roc_ef))),
    caption = sprintf("Red point = Youden-optimal cutoff; pROC direction = %s; n = %d",
                      .roc_ef$direction, nrow(.youden_data))
  ) +
  theme_bw(base_size = 12)
```
